Applicability of SSM and UML for Designing a Search Application for the British Broadcasting Corporation (BBC)

### 2014-11-04 academy-talk.tex
Reviewed each slide at least once now.

@@

% Background
\begin{frame}
  \frametitle{About Me}
  \begin{itemize}
    \pause \item Senior Software Engineer
    \pause \item Future Media: Content Discovery: Search
  \end{itemize}
\end{frame}

@@

% Domain modelling for search -- model business rules as per benefits shown in background, but also keep it generic
\begin{frame}
  \frametitle{Domain Modelling}
  \begin{itemize}
    \item Use cases identify types of content users want to search
    \item Ontology designs exist for specific content domains, e.g. programmes, sport
    \item A search across all content types needs a model to span all domains
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Domain Modelling}
  \framesubtitle{Class Diagram}
  \begin{center}
    \pause \includegraphics[height=0.8\textheight]{plant_model.png}
  \end{center}
\end{frame}

@@

% Discussion and Analysis -- Domain Model -- Polymorphism
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Domain Model}
  \begin{itemize}
    \pause \item Large, fully-inclusive data model useful mainly as communication artefact
    \pause \item Aggregate models hard to keep in sync with changes in respective domains
    \pause \item We can extract a polymorphic supertype containing only properties needed for a search application
    \pause \item We can go further and model a ``Search Result'' facade
  \end{itemize}
\end{frame}

% Discussion and Analysis -- Domain Model -- Duck Typing
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Schemalessness and Duck Typing}
  \begin{itemize}
    \pause \item Statically typed approach: A \emph{Place} always has a ``today's weather forecast'' attribute
    \pause \item Dynamically (``duck'') typed approach: any entity can have a ``today's weather forecast'' attribute and we display this forecast if that attribute is present
    \pause \item Allows other entities in future to have forecasts too (e.g. \emph{Events} such as football matches)
    \pause \item Search application can store all fields from upstream sources and choose when to start using them
  \end{itemize}
\end{frame}

% Discussion -- More SSM: 3 Es?
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Overall Model}
  \begin{itemize}
    \pause \item Promising at a high level
    \pause \item Agile methodology would encourage developing prototype or MVP to test assumptions early
    \pause \item Lacking suitable ``monitoring'' of the system as taught by SSM, e.g. Checkland's ``3 Es'' (efficacy, efficiency, effectiveness)
  \end{itemize}
\end{frame}

% Shortcomings -- Performance? NFRs?
\begin{frame}
  \frametitle{Evaluation and Difficulties}
  \framesubtitle{Non-functional Requirements}
  \begin{itemize}
    \pause \item BBC is a high-traffic website
    \pause \item Content such as news articles are published frequently
    \pause \item How to model requirements around performance, robustness and latency?
  \end{itemize}
\end{frame}

% Shortcomings -- Agile?
\begin{frame}
  \frametitle{Evaluation and Difficulties}
  \framesubtitle{Detailed Design and Agile}
  \begin{itemize}
    \pause \item Models developed at such a high level lack detailed design (not surprisingly)
    \pause \item Agile methodologies encourage designing and delivering a small part of the system early
    \pause \item Detailed design emerges in response to feedback and change
    \pause \item Hard models can be altered in time with these iterations
    \pause \item We can defer outstanding questions about hard models until we might have more information
    \pause \item This seems in line with the continuous feedback cycles within SSM
  \end{itemize}
\end{frame}

% Validation?

% Further Work
\begin{frame}
  \frametitle{Further Work}
  \begin{itemize}
    \pause \item How can the models be effectively validated?
    \pause \item Can the rich picture and use of SSM be repeated and expanded with input from a wider range of experts?
    \pause \item How will the models change when this happens?
    \pause \item Are there more parallels between Agile and SSM that engineers in industry have yet to appreciate?
    \pause \item Should more development processes be considered alongside SSM and UML? e.g. RUP (OpenUP) or AUP?
  \end{itemize}
\end{frame}

>>>>

%gets rid of bottom navigation bars
\setbeamertemplate{footline}[frame number]{}

%gets rid of navigation symbols
\setbeamertemplate{navigation symbols}{}

@@

% Background
\begin{frame}
  \frametitle{About Me}
  \begin{itemize}
    \item Senior Software Engineer
    \item Future Media: Content Discovery: Search
  \end{itemize}
\end{frame}

@@

% Domain modelling for search -- model business rules as per benefits shown in background, but also keep it generic
\begin{frame}
  \frametitle{Domain Modelling}
  \begin{itemize}
    \item Use cases identify types of content users want to search
    \item Things we want to model should emerge from the rich picture
    \item ``TV'' and ``Radio''? Or just ``programmes''?
    \item Ontology designs exist for specific content domains, e.g. programmes, sport
    \item A search across all content types needs a model to span all domains
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Domain Modelling}
  \framesubtitle{Maximal Class Diagram/Domain Model}
  \begin{center}
    \pause \includegraphics[height=0.8\textheight]{plant_model.png}
  \end{center}
\end{frame}

@@

% Discussion and Analysis -- Domain Model -- Polymorphism
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Domain Model}
  \begin{itemize}
    \item Large, maximal data model useful as communication artefact
    \item Aggregate models hard to keep in sync with changes in respective domains
    \item Does Search need to model all of this?
      \begin{itemize}
        \item Polymorphic supertype -- ``Thing'' class?
        \item Facade -- ``SearchResult'' class?
      \end{itemize}
  \end{itemize}
\end{frame}

% Discussion and Analysis -- Domain Model -- Duck Typing
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Schemalessness and Duck Typing}
  \begin{itemize}
    \item Statically typed approach: A \emph{Place} always has a ``today's weather forecast'' attribute
    \item Dynamically (``duck'') typed approach: any entity can have a ``today's weather forecast'' attribute and we display this forecast if that attribute is present
    \item Allows other entities in future to have forecasts too (e.g. \emph{Events} such as football matches)
    \item Search application can store all fields from upstream sources and choose when to start using them
  \end{itemize}
\end{frame}

% Discussion -- More SSM: 3 Es?
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Overall Model: Use cases, behaviour and domain}
  \begin{itemize}
    \item Looks good
    \item Agile? MVP? Test assumptions?
    \item How do we monitor for ``3 Es''?
      \pause \begin{itemize}
        \item Efficacy -- Does the system do what we intended?
        \item Efficiency -- Does it minimise use of resources?
        \item Effectiveness -- Does it fit into our long term goals?
      \end{itemize}
  \end{itemize}
\end{frame}

% Shortcomings -- Agile?
\begin{frame}
  \frametitle{Evaluation and Difficulties}
  \framesubtitle{Detailed Design and Agile}
  \begin{itemize}
    \item High-level design lacks detail (not surprisingly)
    \item Agile encourages delivering something tangible early
    \item Shouldn't the full detailed design emerge iteratively?
    \item Can we iteratively build up and maintain hard models?
    \item Can we defer pending questions in our hard models until we know more?
    \item SSM encourages continuous feedback cycles -- does this align with Agile?
    \item Bustard, Dave, and Frank Keenan. ``Soft Systems Methodology: An Aid to Agile Development?.''
      In \emph{Information Systems Development}, pp. 25-38. Springer US, 2009.
  \end{itemize}
\end{frame}

% Shortcomings -- Performance? NFRs?
\begin{frame}
  \frametitle{Evaluation and Difficulties}
  \framesubtitle{Non-functional Requirements}
  \begin{itemize}
    \item Performance
    \item Availability/uptime
    \item Latency between publishing and searchability
    \item Secure/robust
    \item Maintainability -- the holistic system includes developers!
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Further Work}
  \begin{itemize}
    \item Can SSM take a more formal place in our processes?
    \item Are there more parallels between Agile and SSM that engineers in industry have yet to appreciate?
    \item Can we streamline modelling and design to fit an iterative process?
  \end{itemize}
\end{frame}
